Personality and political beliefs across the lifespan

### 2019-11-23 scripts/manuscript.Rmd
one comment for improving efficiency (slightly)
library(grid) to provide space between facet_wrap, spacing the x-axis

```{r setup, include = FALSE}
library(papaja)
library(rio)
library(here)
library(tidyverse)
library(psych)

@@

#filter out those who took the survey too quickly (the time cut offs were predetermined and based on which condition the participant was randomly assigned to)
data <- data %>% 
  pivot_longer(cols = Video_DO_V4:Video_DO_V1, 
               names_to = "Video_DO", 
               values_to = "Value",
               names_prefix = "Video_DO_",
               values_drop_na = TRUE) %>% 
  filter(Video_DO == 'V1' & Duration__in_seconds_ >316 | 
           Video_DO == 'V2' & Duration__in_seconds_ > 340 | 
           Video_DO == 'V3' & Duration__in_seconds_ > 398 |
           Video_DO == 'V4' & Duration__in_seconds_ > 666 | 
           Video_DO == 'V5' & Duration__in_seconds_ > 565 | 
           Video_DO == 'V6' & Duration__in_seconds_ > 538) %>% 
  filter(Duration__in_seconds_ < 5000) #this person spent too long taking the survey
```

@@

#econonmic conservatism subscale 
scaleKey <- c(0, 1, 0, 0, -1, 1, 0, 0, 1, 1, 0, 0)
economic <- scoreItems(keys = scaleKey, items = data[items], 
  totals = FALSE, missing = FALSE, min = 0, max = 100)
data <- data %>% 
  mutate(SEC_economic_conservatism = economic$scores[,1])

#social conservatism subscale
scaleKey <- c(-1, 0, 1, 1, 0, 0, 1, 1, 0, 0, 1, 1)
social <- scoreItems(keys = scaleKey, items = data[items], 
  totals = FALSE, missing = FALSE, min = 0, max = 100)
data <- data %>% 
  mutate(SEC_social_conservatism = social$scores[,1])
```

@@

#Extraversion
scaleKey <- c(1, 0, 0, 0, 0, -1, 0, 0, 0, 0)
extraversion <- scoreItems(keys = scaleKey, items = data[items], 
  totals = FALSE, missing = FALSE, min = 1, max = 7)
data <- data %>% 
  mutate(Extraversion = extraversion$scores[,1])

#Agreeableness
scaleKey <- c(0, -1, 0, 0, 0, 0, 1, 0, 0, 0)
agree <- scoreItems(keys = scaleKey, items = data[items], 
  totals = FALSE, missing = FALSE, min = 1, max = 7)
data <- data %>% 
  mutate(Agreeableness = agree$scores[,1])

#Conscientiousness
scaleKey <- c(0, 0, 1, 0, 0, 0, 0, -1, 0, 0)
consc <- scoreItems(keys = scaleKey, items = data[items], 
  totals = FALSE, missing = FALSE, min = 1, max = 7)
data <- data %>% mutate(Conscientiousness = consc$scores[,1])

#Emotional stability
scaleKey <- c(0, 0, 0, -1, 0, 0, 0, 0, 1, 0)
emo <- scoreItems(keys = scaleKey, items = data[items], 
  totals = FALSE, missing = FALSE, min = 1, max = 7)
data <- data %>% 
  mutate(Emotional_stability = emo$scores[,1])

#Openness to experiences
scaleKey <- c(0, 0, 0, 0, 5, 0, 0, 0, 0, -1)
open <- scoreItems(keys = scaleKey, items = data[items], 
  totals = FALSE, missing = FALSE, min = 1, max = 7)
data <- data %>% 
  mutate(Openness_to_experiences = open$scores[,1])
```

@@

```{r tidying}
#last step is to tidy our data
tidy <- data %>% 
  pivot_longer(cols = c(sec_economic_conservatism, sec_social_conservatism),
               names_to = "conservatism_type",
               names_prefix = "sec_",
               values_to = "conservatism_score") %>% 
  mutate(conservatism_type = recode(conservatism_type, `economic_conservatism` = "Economic conservatism",
                                    `social_conservatism` = "Social conservatism")) %>%
  pivot_longer(cols = agreeableness:emotional_stability,
               names_to = "personality_trait",
               values_to = "personality_score") %>% 
  mutate(personality_trait = recode(personality_trait, `agreeableness` = "Agreeableness",
                                    `conscientiousness` = "Conscientiousness",
                                    `openness_to_experiences` = "Openness to experiences",
                                    `extraversion` = "Extraversion",
                                    `emotional_stability` = "Emotional stability"))
```

@@

apa_table(tidy_table, 
            caption = "Mean Personality Trait Scores Across the Lifespan",
          col_spanners = list(`Age Group` = c(2, 6)),
          note = "Personality trait scores reported on a 1-7 scale."
)
```

@@

ggplot(tidy, aes(x = age, y = personality_score, color = personality_trait)) +
  geom_smooth(se=FALSE, method = "lm")+
  labs(title = "Personality traits across the lifespan", subtitle = "Big 5 traits self-reported by adults aged 20-68",
       x= "Age", y = "How strongly participants identify with the personality trait", 
       color = "Personality Trait")+
  theme_minimal()+
  scale_color_viridis_d()

@@

ggplot(tidy, aes(x = age, y = conservatism_score, color = conservatism_type)) +
  geom_smooth(se=FALSE, method = "lm")+
  scale_color_viridis_d() +
  labs(title = "Conservatism across the lifespan (adults aged 20-68)", subtitle = "Difference in Economic and Social Conservatism by Gender",
       x= "Age", y = "Conservatism Score (higher = more conservative)",
       color = "Conservatism Type")+
  theme_minimal()+
  facet_wrap(~gender)
```

```{r messy_data}
messy <- tidy %>% 
  pivot_wider(
    names_from = conservatism_type, 
    values_from = conservatism_score) %>%
   pivot_wider(
     names_from = personality_trait,
     values_from = personality_score)

#we can use this messy data to run lm models
```

>>>>

```{r setup, include = FALSE}
library(papaja)
library(rio)
library(here)
library(tidyverse)
library(psych)
#install.packages("ggsave")
library(grid)

@@

#filter out those who took the survey too quickly (the time cut offs were predetermined and based on which condition the participant was randomly assigned to)
data <- data %>% 
  pivot_longer(cols = Video_DO_V4:Video_DO_V1,
               names_to = "Video_DO",
               values_to = "Value",
               names_prefix = "Video_DO_",
               values_drop_na = TRUE) %>%
  filter(Video_DO == 'V1' & Duration__in_seconds_ > 316 |
           Video_DO == 'V2' & Duration__in_seconds_ > 340 |
           Video_DO == 'V3' & Duration__in_seconds_ > 398 |
           Video_DO == 'V4' & Duration__in_seconds_ > 666 |
           Video_DO == 'V5' & Duration__in_seconds_ > 565 |
           Video_DO == 'V6' & Duration__in_seconds_ > 538) %>%
  filter(Duration__in_seconds_ < 5000) #this person spent too long taking the survey
```

@@

#econonmic conservatism subscale 
scaleKey <- c(0, 1, 0, 0, -1, 1, 0, 0, 1, 1, 0, 0)
economic <- scoreItems(
  keys = scaleKey,
  items = data[items],
  totals = FALSE,
  missing = FALSE,
  min = 0, max = 100)
data <- data %>% 
  mutate(SEC_economic_conservatism = economic$scores[,1])

#social conservatism subscale
scaleKey <- c(-1, 0, 1, 1, 0, 0, 1, 1, 0, 0, 1, 1)
social <- scoreItems(
  keys = scaleKey,
  items = data[items],
  totals = FALSE,
  missing = FALSE,
  min = 0, 
  max = 100)
data <- data %>% 
  mutate(SEC_social_conservatism = social$scores[,1])
```

@@

#Extraversion
scaleKey <- c(1, 0, 0, 0, 0, -1, 0, 0, 0, 0)
extraversion <- scoreItems(
  keys = scaleKey,
  items = data[items],
  totals = FALSE,
  missing = FALSE,
  min = 1, max = 7)
data <- data %>% 
  mutate(Extraversion = extraversion$scores[,1])

#Agreeableness
scaleKey <- c(0, -1, 0, 0, 0, 0, 1, 0, 0, 0)
agree <- scoreItems(
  keys = scaleKey,
  items = data[items],
  totals = FALSE,
  missing = FALSE,
  min = 1,
  max = 7)
data <- data %>%
  mutate(Agreeableness = agree$scores[,1])

#Conscientiousness
scaleKey <- c(0, 0, 1, 0, 0, 0, 0, -1, 0, 0)
consc <- scoreItems(
  keys = scaleKey,
  items = data[items],
  totals = FALSE,
  missing = FALSE,
  min = 1,
  max = 7)
data <- data %>%
  mutate(Conscientiousness = consc$scores[,1])

#Emotional stability
scaleKey <- c(0, 0, 0, -1, 0, 0, 0, 0, 1, 0)
emo <- scoreItems(
  keys = scaleKey,
  items = data[items],
  totals = FALSE,
  missing = FALSE,
  min = 1,
  max = 7)
data <- data %>%
  mutate(Emotional_stability = emo$scores[,1])

#Openness to experiences
scaleKey <- c(0, 0, 0, 0, 5, 0, 0, 0, 0, -1)
open <- scoreItems(
  keys = scaleKey,
  items = data[items],
  totals = FALSE,
  missing = FALSE,
  min = 1, max = 7)
data <- data %>%
  mutate(Openness_to_experiences = open$scores[,1])
```

@@

```{r tidying}
#last step is to tidy our data
tidy <- data %>%
  pivot_longer(cols = c(
    sec_economic_conservatism,
    sec_social_conservatism),
    names_to = "conservatism_type",
    names_prefix = "sec_",
    values_to = "conservatism_score") %>% 
  mutate(conservatism_type = recode(
    conservatism_type, 
    `economic_conservatism` = "Economic conservatism",
    `social_conservatism` = "Social conservatism")) %>%
  pivot_longer(cols = agreeableness:emotional_stability,
               names_to = "personality_trait",
               values_to = "personality_score") %>%
  mutate(personality_trait =
           recode(personality_trait,
                  `agreeableness` = "Agreeableness",
                  `conscientiousness` = "Conscientiousness",
                  `openness_to_experiences` = "Openness to experiences",
                  `extraversion` = "Extraversion",
                  `emotional_stability` = "Emotional stability"))
```

@@

apa_table(tidy_table,
          caption = "Mean Personality Trait Scores Across the Lifespan",
          col_spanners = list(`Age Group` = c(2, 6)),
          note = "Personality trait scores reported on a 1-7 scale.")
```

@@

ggplot(tidy, aes(x = age, y = personality_score, color = personality_trait)) +
  geom_smooth(se = FALSE, method = "lm") +
  labs(
    title = "Personality traits across the lifespan",
    subtitle = "Big 5 traits self-reported by adults aged 20-68",
    x = "Age", y = "How strongly participants identify with the personality trait",
    color = "Personality Trait") +
  theme_minimal() +
  scale_color_viridis_d()

@@

ggplot(tidy, aes(x = age, y = conservatism_score, color = conservatism_type)) +
  geom_smooth(se=FALSE, method = "lm") +
  scale_color_viridis_d() +
  labs(
    title = "Conservatism across the lifespan (adults aged 20-68)",
    subtitle = "Difference in Economic and Social Conservatism by Gender",
    x = "Age", y = "Conservatism Score (higher = more conservative)",
    color = "Conservatism Type") +
  theme_minimal() +
  theme(panel.spacing = unit(2, "lines")) +
  facet_wrap(~gender)
```

```{r messy_data}
messy <- tidy %>% 
  pivot_wider(
    names_from = conservatism_type, 
    values_from = conservatism_score) %>%
  pivot_wider(
    names_from = personality_trait,
    values_from = personality_score)
#we can use this messy data to run lm models
```
